Manuscript Figures for Genome Biology

### 2018-05-11 manuscript_figures/manuscript_figures.Rmd
Refine method rank heatmap; add case-study specific heatmaps
Method ranking heatmap is now a standalone function so it can
be reused to make average ranks across case studies, datsets,
and/or covariates. The default now colors by mean proportion of
maximum rejections instead of mean rank.

# assumes sb objects are in the following location, which contains subfolders
# for each casestudy (if this isn't true, then parsing the case study and 
# dataset names later on will be incorrect)
path <- "../sb-objects"

@@

# point to where yeast results are (within 'path' defined above)
resdir <- file.path(path, "RNAseq/yeast-results")

@@

```{r Figure2}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
resfile <- file.path(resdir, "de5.rds")
de5 <- readRDS(file=resfile)
res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))

@@

```{r, Figure 4}
library(SummarizedBenchmark)
library(stringr)
library(viridis)

@@

# remove simulation objects
sim <- unique(c(which(grepl("simulations", objects)),
               which(grepl("yeast", objects)),
               which(grepl("polyester", objects))))

if (length(sim) > 0){
  objects <- objects[-sim]
}

# create tidy data frame where each row is a method / dataset observation
# of a rank 
ranks <- data.frame()
obj.names <- str_split(objects, pattern = "/")
casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
dataset <- unlist(lapply(obj.names, function(x) x[4] ))

for (i in seq_along(objects)){
  x <- readRDS(objects[i])
  assayNames(x) <- "qvalue"
  x <- addDefaultMetrics( x )

  hasResults <- apply(!is.na( assays(x)[["qvalue"]] ), 2, all )

  if (sum(hasResults) > 0){
    tmp <- estimatePerformanceMetrics(x, alpha, tidy=TRUE) %>%
      filter( performanceMetric == "rejections") %>%
      dplyr::rename( method = blabel) %>%
      filter( is.na(param.alpha) | (param.alpha == alpha)) %>%
      filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
      filter( !method == "unadjusted") %>%
      select( method, value ) %>%
      dplyr::rename( nrejects = value) %>%
      mutate( method = gsub("-df03", "", method)) %>%
      mutate( method = gsub("-a10", "", method)) %>%
      na.omit() %>%
      mutate( rank = rank(nrejects) / n() ) %>%
      mutate(casestudy = casestudy[i],
             dataset = dataset[i])

    ranks <- rbind(ranks, tmp)
  }
}

# average over datasets within case study
ranks_avg <- ranks %>% 
  group_by( casestudy, method) %>%
  summarize( mean_rank = mean(rank, na.rm = TRUE))

case_dat <- ranks_avg %>%
  group_by( casestudy) %>%
  summarize( nmethods = n())

method_dat <- ranks_avg %>%
  group_by( method) %>%
  summarize( meta_rank = mean(mean_rank))

ranks_avg <- left_join(ranks_avg, method_dat, by ="method")
ranks_avg <- left_join(ranks_avg, case_dat, by ="casestudy")

# reorder rows and columns
ranks_avg$casestudy <- factor(ranks_avg$casestudy, 
                              levels=unique(ranks_avg$casestudy[order(ranks_avg$nmethods)]))
ranks_avg$method <- factor(ranks_avg$method,
                           levels=unique(ranks_avg$method[order(ranks_avg$meta_rank)]))

# heatmap : rows method, cols casestudy
Fig4 <- ggplot(ranks_avg, aes(x = casestudy, y = method, fill = mean_rank)) + 
  geom_raster() + 
  scale_fill_viridis() +
  theme_bw() +
  xlab("Case Study") +
  ylab("Method") +
  labs(fill = "Mean rank")

Fig4
ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4, dev = pdf)
```

@@

>>>>

# assumes sb objects are in the following location, which contains subfolders
# for each casestudy (if this isn't true, then parsing the case study and 
# dataset names later on will be incorrect)
path <- "../results"

@@

@@

```{r Figure2}
objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("yeast", objects) & grepl("de5", objects)] 
de5 <- readRDS(file=resfile)
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))

@@

```{r, Figure 4}
library(SummarizedBenchmark)
library(stringr)

@@

# remove simulation objects
sim <- unique(c(which(grepl("simulations", objects)),
               which(grepl("yeast", objects)),
               which(grepl("polyester", objects))))
if (length(sim) > 0){
  objects <- objects[-sim]
}

# remove microbiome objects from 'null' comparisons
# removing baxter, goodrich and papa OTU-level (since they mostly found 
# nothing for all methods and the genus-level analysis is also present)
null <- which(grepl(c("baxter|goodrich|papa"), objects) & 
              grepl(c("otu"), objects))

if (length(null) > 0){
  objects <- objects[-null]
}

# remove covariates from individual cases studies that weren't informative
## TODO ##

# grab casestudy names
obj.names <- str_split(objects, pattern = "/")
casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
dataset <- unlist(lapply(obj.names, function(x) x[4] ))

Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10)
Fig4
ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4)

# heatmap : rows method, cols casestudy- one color scale 
Fig4alt <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10, 
                        colorLow = "white", colorHigh = "blue", 
                        colorNA = "grey50")
Fig4alt
ggsave(file.path(outdir, "Figure4alt.pdf"), width=6, height=4)

##
## AR: alternative suggestion to heatmap, feel free to discard! ##
##

Fig4b <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
                         linePlot = TRUE)
Fig4b
```

@@

# Supplementary Figures

## Within case-study heatmaps of method rankings

Here we generate heatmaps for each case study, where each column represents a different
dataset. Here we pick one representative covariate and method (if applicable) 
for each case study. See the individual Rmds for more detailed plots comparing, e.g.
different covariates on the same dataset.

```{r}
objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )

# grab casestudy names
obj.names <- str_split(objects, pattern = "/")
casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
dataset <- unlist(lapply(obj.names, function(x) x[4] ))
```

### ChIP-seq

```{r}
wcs <- which(grepl("ChIPseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
  ggtitle("ChIP-Seq")
```

### GSEA

```{r}
#by dataset
wcs <- which(grepl("GSEA", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
  ggtitle("GSEA")
```

### GWAS

```{r}
# by covariate
wcs <- which(grepl("GWAS", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
                      function(x) x[2]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate") + 
  ggtitle("GWAS")
```

### Microbiome

```{r}
# by dataset, excluding otu when genus is available
# excluding log-ubiquity
# excluding enigma-so4 (no discoveries)
wcs <- which(grepl("microbiome", casestudy) & 
              !(grepl(c("baxter|goodrich|papa"), dataset) & 
              grepl(c("otu"), dataset)) &
              !grepl("log-ubiquity", dataset) &
              !grepl("so4", dataset))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
  ggtitle("Microbiome")
```

### RNA-seq
```{r}
# by dataset
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("polyester|yeast"), dataset) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
  ggtitle("RNA-seq")
```

### scRNA-seq

```{r}
wcs <- which(grepl("scRNAseq", casestudy))
objs <- objects[wcs]
ds <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark-"), 
                      function(x) x[1]))
m0 <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark-"), 
                      function(x) x[2]))
meth <- unlist(lapply(str_split(m0, pattern = "-"), 
                      function(x) x[1]))
cv <- gsub(".rds", "", unlist(lapply(str_split(m0, pattern = "-"), 
                      function(x) x[2])))

# by dataset
plotMethodRanks(objs, colLabels = ds, alpha = 0.10, xlab = "Dataset") + 
  ggtitle("scRNA-seq")

# by method
plotMethodRanks(objs, colLabels = meth, alpha = 0.10, xlab = "Method") + 
  ggtitle("scRNA-seq")

# by covariate
plotMethodRanks(objs, colLabels = cv, alpha = 0.10, xlab = "Covariate") + 
  ggtitle("scRNA-seq")
```

### Yeast Simulation

```{r}
TODO
```

### Polyester Simulation

```{r}
TODO
```
